---
title: "Adjusted sample size using legacy site buffer"
author: Willian Vieira
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
        toc_depth: 5
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
EVAL = TRUE
```

## Setup

```{r libraries,warning=F,message=F}
library(tidyverse)
library(ggpubr)
library(kableExtra)
library(gridtext)
library(ggtext)
library(sf)
library(spatstat)
library(spsurvey)
library(MBHdesign)

set.seed(0.0)

hexas <- readRDS('../data/hexa_complete.RDS')

# First filter for hexagons that have at least 20% of sampleable pixel habitats (non NAs)
hexas <- subset(hexas, propNA <= 0.8)

# calculate p
hexas$p <- (hexas$hab_prob * hexas$cost_prob) / sum(hexas$hab_prob * hexas$cost_prob)

# remove zeros
hexas <- subset(hexas, p != 0)
```

```{r define ecoregions, echo=F, eval = EVAL}
# filter for ecoregions with at least 10 ecoregion
eco_sim <- hexas %>%
    st_drop_geometry() %>%
    group_by(ecoregion) %>%
    summarise(nbLegacy = sum(legacySite > 0)) %>%
    filter(nbLegacy > 10) %>%
    pull(ecoregion)
```


## Adjust sample size in function of buffer size

```{r sim, warning=FALSE, eval = EVAL}
sampleProportion <- 0.02 # 2% of available hexagons

buffer_sizes <- seq(0, 35000, 2500)
legacy_minimum <- c(0, 2, 4)

sampleSize <- data.frame()
for(eco in eco_sim)
{
    hexa_eco <- hexas %>%
        filter(ecoregion == eco) %>%
        st_centroid()

    # number of available hexagons to sample
    nb_hexas <- nrow(hexa_eco)
    
    for(lc_min in legacy_minimum)
    {
        hexa_legacy <- hexas %>%
            filter(ecoregion == eco & legacySite > lc_min)

        # initial sample size
        init_ss <- nb_hexas * sampleProportion

        for(bf_sz in buffer_sizes)
        {
            # get buffer around legacy sites
            legacy_buffer <- hexa_legacy %>%
                st_buffer(bf_sz) %>%
                st_union()
            
            adj_ss <- hexa_eco %>%
                st_intersects(legacy_buffer) %>%
                unlist() %>%
                sum() * sampleProportion

            sampleSize <- rbind(
                sampleSize,
                data.frame(
                    ecoregion = eco,
                    legacy_min = lc_min,
                    bufferSize = bf_sz/1000,
                    n_adj = init_ss - adj_ss
                )
            )            
        }
    }
}
```

```{r fig1, echo=FALSE,fig.align='center',fig.width=10,fig.height=9, eval = EVAL}
sampleSize %>%
    mutate(legacy_min = as.factor(legacy_min)) %>%
    ggplot(aes(bufferSize, n_adj, color = legacy_min)) +
        geom_line() +
        facet_wrap(~ecoregion, scales = 'free_y', ncol = 4) +
        xlab('Buffer size (Km)') +
        ylab('Adjusted sample size') +
        theme_classic() +
        theme(
            legend.position = 'bottom'
        ) +
        ggtitle('Adjusted sample size after removing hexagons within the buffer around legacy sites') 
```


## GRTS simulation

```{r,echo=FALSE,eval = EVAL}
# Run GRTS for four different sets
buffer_sizes <- c(0, 5000, 10000, 20000)

# round adjusted N
sampleSize <- sampleSize %>%
    mutate(
        N = round(n_adj, 0),
        bufferSize_km = bufferSize * 1000
    )

# prepare sample frame
sampleFrame_legacy <- hexas %>%
    filter(ecoregion %in% eco_sim) %>%
    mutate(
        eco_name = paste0('eco_', ecoregion), # to match design name
        legacy_var = ifelse(legacySite == 0, NA, legacySite),
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )


# coordinates in matrix format for MBHdesign
coord_mt <- sampleFrame_legacy %>%
    st_coordinates()

legacySites <- sampleFrame_legacy %>%
    filter(legacySite > 0) %>%
    st_coordinates()

# adjust inclusion probability
for(sgm in 1:length(buffer_sizes))
{
    # inclusion probability weighted by N
    sampleFrame_legacy[[paste0('mdcaty_', sgm)]] <-
        sum(subset(sampleSize, bufferSize_km == buffer_sizes[sgm] & legacy_min == 0)$N) * sampleFrame_legacy$p/sum(sampleFrame_legacy$p)

    # adjusted inclusion probability in function legacy site and buffer size
    sampleFrame_legacy[[paste0('pAdj_', sgm)]] <- 
        MBHdesign::alterInclProbs(
            legacy.sites = legacySites,
            potential.sites = coord_mt,
            inclusion.probs = sampleFrame_legacy[[paste0('mdcaty_', sgm)]],
            sigma = buffer_sizes[sgm]
        )
}

# for buffer size == 0, incl prob does not change
sampleFrame_legacy$pAdj_1 <- sampleFrame_legacy$mdcaty_1



# run GRTS
out_sim <- list()
for(sgm in 1:length(buffer_sizes))
{
    # Sample size by stratum (ecoregion)
    Stratdsgn_legacy  <- c()
    for(eco in eco_sim)
        Stratdsgn_legacy <- append(
            Stratdsgn_legacy,
            setNames(
                subset(sampleSize, legacy_min == 0 & ecoregion == eco & bufferSize_km == buffer_sizes[sgm])$N,
                paste0('eco_', eco)
            )
        )

    # remove ecoregions with sample size == 0
    Stratdsgn_legacy <- Stratdsgn_legacy[Stratdsgn_legacy > 0]
    
    out_sample_legacy <- spsurvey::grts(
        sframe = sampleFrame_legacy,
        n_base = Stratdsgn_legacy,
        stratum_var = 'eco_name',
        aux_var = paste0('pAdj_', sgm)
    )

    out_sim[[sgm]] <- out_sample_legacy$sites_base$ET_Index
}
```


### GRTS simulation for divided ecoregions (1 replication)

Divided ecoregions with at least 10 legacy sites: 100, 101, 103, 217, and 96.

For each ecoregion, we can visualize the selected sites from the GRTS (blue) in function of the legacy sites (red) and the buffer size (sigma; ligth red).
Note that `sigma = 0` means that neither the inclusion probability nor the sample size (N) was adjusted.

Second figure shows how the habitat types are proportionally distributed between (i) all the hexagons, (ii) only the legacy hexagons, and the selected hexagons depending on the value of sigma.


```{r,echo=FALSE,fig.align='center',fig.width=12,fig.height=18,warning=FALSE}
habProp_f <- function(DT)
    DT %>%
        st_drop_geometry() %>%
        select(starts_with('land_ca_')) %>%
        summarise(across(everything(), sum, na.rm = TRUE)) %>%
        pivot_longer(everything(), names_to = 'land')


habitat_colors <- c(land_ca_1 = rgb(0, 61, 0, maxColorValue = 255),
                    land_ca_2 = rgb(148, 156, 112, maxColorValue = 255),
                    land_ca_5 = rgb(20, 140, 61, maxColorValue = 255),
                    land_ca_6 = rgb(92, 117, 43, maxColorValue = 255),
                    land_ca_8 = rgb(179, 138, 51, maxColorValue = 255),
                    land_ca_10 = rgb(225, 207, 138, maxColorValue = 255),
                    land_ca_11 = rgb(156, 117, 84, maxColorValue = 255),
                    land_ca_12 = rgb(186, 212, 143, maxColorValue = 255),
                    land_ca_13 = rgb(64, 138, 112, maxColorValue = 255),
                    land_ca_14 = rgb(107, 163, 138, maxColorValue = 255),
                    land_ca_16 = rgb(168, 171, 174, maxColorValue = 255),
                    land_ca_21 = '#03045e',
                    land_ca_22 = '#023e8a',
                    land_ca_23 = '#0077b6',
                    land_ca_24 = '#0096c7',
                    land_ca_25 = '#00b4d8',
                    land_ca_26 = '#48cae4',
                    land_ca_27 = '#90e0ef',
                    land_ca_28 = '#ade8f4',
                    land_ca_29 = '#caf0f8',
                    land_ca_30 = '#E6FFFF')

landNames = setNames(
    c(
        'All hexagons',
        'Legacy sites',
        paste0('Sigma = ', buffer_sizes)
    ),
    1:(length(buffer_sizes) + 2)
)
        

# keep only divided ecoregions
eco_sim_div <- grep('N|S', eco_sim, value = TRUE)

for(eco in eco_sim_div)
{
    legacy <- hexas %>%
        filter(ecoregion == eco & legacySite > 0)
    
    plotList <- list()
    sel_ls <- list()
    for(sgm in 1:length(buffer_sizes))
    {
        legacy_buffer <- legacy %>%
            st_buffer(dist = buffer_sizes[sgm]) %>%
            st_union()
        
        sel_ls[[sgm]] <- hexas %>%
            filter(ecoregion == eco & ET_Index %in% out_sim[[sgm]])

        plotList[[sgm]] <-
            ggplot() +
                geom_sf(data = st_union(subset(hexas, ecoregion == eco)[, 1])) +
                geom_sf(data = legacy_buffer, fill = rgb(.85, 0, 0, 0.05), color = 'transparent') +
                geom_sf(data = sel_ls[[sgm]], fill = 'darkblue', color = 'darkblue') +
                geom_sf(data = legacy, fill = rgb(1, 0, 0, 0.4), col = 'transparent') +
                ggtitle(
                    paste0(
                        'Sigma = ', buffer_sizes[sgm],
                        ';  N = ', subset(sampleSize, legacy_min ==0 & ecoregion == eco & bufferSize_km == buffer_sizes[sgm])$N
                        )
                    )
    }
    title_text <- paste0(
        '<br><br>**<span style="font-size:14pt;">Ecoregion = ', eco, '</span>**<br>',
        '# legacy sites = ', nrow(legacy)        
    )
    p1 <- ggarrange(plotlist = plotList, ncol = 2, nrow = 2) %>%
            annotate_figure(
                top = gridtext::richtext_grob(title_text)
                )


    # get habitat proportion
    land_prop <- map_dfr(
        append(
            list(
                subset(hexas, ecoregion == eco),
                legacy
            ),
            sel_ls
        ),
        habProp_f,
        .id = 'sim'
    )
    
    p2 <- ggplot(land_prop, aes(fill = land, x = sim, y = value)) +
            geom_bar(position = 'fill', stat = 'identity') +
            scale_fill_manual(values = habitat_colors) +
            theme_classic() +
            ylab('Habitat proportion') + 
            xlab('') +
            scale_x_discrete(labels = landNames)
    
    # plot
    print(ggarrange(
        p1,
        p2,
        ncol = 1, nrow = 2,
        heights = c(2, 0.7)
    ))
}
```


### GRTS simulation for entire ecoregions (1 replication)

Entire ecoregions with at least 10 legacy sites: 100, 101, 102, 103, 117, 217, 72, 74, 75, 96, 9.

```{r,echo=FALSE,fig.align='center',fig.width=12,fig.height=18,warning=FALSE}
# keep only divided ecoregions
for(eco in setdiff(eco_sim, eco_sim_div))
{
    legacy <- hexas %>%
        filter(ecoregion == eco & legacySite > 0)
    
    plotList <- list()
    sel_ls <- list()
    for(sgm in 1:length(buffer_sizes))
    {
        legacy_buffer <- legacy %>%
            st_buffer(dist = buffer_sizes[sgm]) %>%
            st_union()
        
        sel_ls[[sgm]] <- hexas %>%
            filter(ecoregion == eco & ET_Index %in% out_sim[[sgm]])

        plotList[[sgm]] <-
            ggplot() +
                geom_sf(data = st_union(subset(hexas, ecoregion == eco)[, 1])) +
                geom_sf(data = legacy_buffer, fill = rgb(.85, 0, 0, 0.05), color = 'transparent') +
                geom_sf(data = sel_ls[[sgm]], fill = 'darkblue', color = 'darkblue') +
                geom_sf(data = legacy, fill = rgb(1, 0, 0, 0.4), col = 'transparent') +
                ggtitle(
                    paste0(
                        'Sigma = ', buffer_sizes[sgm],
                        ';  N = ', subset(sampleSize, legacy_min ==0 & ecoregion == eco & bufferSize_km == buffer_sizes[sgm])$N
                        )
                    )
    }
    title_text <- paste0(
        '<br><br>**<span style="font-size:14pt;">Ecoregion = ', eco, '</span>**<br>',
        '# legacy sites = ', nrow(legacy)        
    )
    p1 <- ggarrange(plotlist = plotList, ncol = 2, nrow = 2) %>%
            annotate_figure(
                top = gridtext::richtext_grob(title_text)
                )


    # get habitat proportion
    land_prop <- map_dfr(
        append(
            list(
                subset(hexas, ecoregion == eco),
                legacy
            ),
            sel_ls
        ),
        habProp_f,
        .id = 'sim'
    )
    
    p2 <- ggplot(land_prop, aes(fill = land, x = sim, y = value)) +
            geom_bar(position = 'fill', stat = 'identity') +
            scale_fill_manual(values = habitat_colors) +
            theme_classic() +
            ylab('Habitat proportion') + 
            xlab('') +
            scale_x_discrete(labels = landNames)
    
    # plot
    print(ggarrange(
        p1,
        p2,
        ncol = 1, nrow = 2,
        heights = c(2, 0.7)
    ))
}
```



## What would be the optimum `sigma` for the subsequent years?

If we set the target to 2% of available hexagons, but could sample only half of it in the first year, how could we use to above approach to difine the remaining N for the next year?

### Run GRTS for year 1

```{r}
# keep ecoregions with at least 300 available hexagons (N = 6)
eco_sim <- hexas %>%
    st_drop_geometry() %>%
    filter(ecoregion %in% grep('N|S', ecoregion, value = TRUE, invert = TRUE)) %>%
    group_by(ecoregion) %>%
    summarise(n = n()) %>% 
    filter(n > 300) %>%
    pull(ecoregion)


# Sample size by stratum (ecoregion)
Stratdsgn  <- c()
for(eco in eco_sim)
    Stratdsgn <- append(
        Stratdsgn,
        setNames(
            round(nrow(subset(hexas, ecoregion == eco)) * 0.02, 0),
            paste0('eco_', eco)
        )
    )


# prepare sample frame
sampleFrame <- hexas %>%
    filter(ecoregion %in% eco_sim) %>%
    mutate(
        eco_name = paste0('eco_', ecoregion), # to match design name
        mdcaty  = sum(Stratdsgn) * p/sum(p),
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )
    
# run GRTS    
out_sample <- spsurvey::grts(
    sframe = sampleFrame,
    n_base = Stratdsgn,
    stratum_var = 'eco_name',
    aux_var = 'mdcaty'
)
```

Now we assume we can only sample half (1%) of the selected hexagons this year:

```{r}
generate_v <- function(n) {
    out = rep(0, n)
    out[sample(1:n, round(n/2, 0))] = 1
    return(as.factor(out))
}

# create a new column with 0-1 sampled rows
selected_hexas <- hexas %>%
    filter(ET_Index %in% out_sample$sites_base$ET_Index) %>%
    group_by(ecoregion) %>%
    mutate(sampled = generate_v(n()))
```

What would be the optimum buffer size so the adjusted N is closer to half of the sample size?

```{r,echo=FALSE,warning=FALSE}
buffer_sizes <- seq(13000, 36000, 300)

sampleSize <- data.frame()
for(eco in eco_sim)
{
    hexa_eco <- hexas %>%
        filter(ecoregion == eco) %>%
        st_centroid()

    # number of available hexagons to sample
    nb_hexas <- nrow(hexa_eco)

    hexa_selected <- selected_hexas %>%
        filter(ecoregion == eco & sampled == 1)

    # initial sample size
    init_ss <- Stratdsgn[paste0('eco_', eco)]

    for(bf_sz in buffer_sizes)
    {
        # get buffer around legacy sites
        legacy_buffer <- hexa_selected %>%
            st_buffer(bf_sz) %>%
            st_union()
        
        adj_ss <- hexa_eco %>%
            st_intersects(legacy_buffer) %>%
            unlist() %>%
            sum() * 0.02

        sampleSize <- rbind(
            sampleSize,
            data.frame(
                ecoregion = eco,
                bufferSize = bf_sz/1000,
                n_adj = init_ss - adj_ss
            )
        )            
    }
}


selected_N <- selected_hexas %>%
    st_drop_geometry() %>%
    filter(sampled == 1) %>%
    group_by(ecoregion) %>%
    summarise(n = n())

best_bufferSize <- data.frame()
for(eco in eco_sim)
{
    expected_n <- unname(Stratdsgn[paste0('eco_', eco)] - subset(selected_N, ecoregion == eco)$n)

    better_bs <- subset(sampleSize, ecoregion == eco)$bufferSize[which.min(abs(expected_n - subset(sampleSize, ecoregion == eco)$n_adj))]

    best_bufferSize <- rbind(
        best_bufferSize,
        data.frame(
            ecoregion = eco,
            exp_n = expected_n,
            best_bs = better_bs
        )
    )
}

best_bufferSize %>%
    mutate(
        Ecoregion = ecoregion,
        `Optimum buffer size (Km)` = best_bs
    ) %>%
    select(4:5) %>%
    kbl() %>%
        kable_paper("hover", full_width = F)

```

### Run GRTS for year 2

The sampled hexagons (1%) of year 1 now become the legacy site for the year 2. We will use the average of the optimum buffer size among the ecoregions (`r round(mean(best_bufferSize$best_bs), 2)` Km) to reduce `N` and concentrate new hexagons in regions not yet covered.

```{r, echo=FALSE,warning=FALSE}
# Sample size by stratum (ecoregion)
Stratdsgn  <- c()
for(eco in eco_sim)
    Stratdsgn <- append(
        Stratdsgn,
        setNames(
            round(subset(sampleSize, ecoregion == eco & bufferSize == subset(best_bufferSize, ecoregion == eco)$best_bs)$n_adj, 0),
            paste0('eco_', eco)
        )
    )

# prepare sample frame
sampleFrame <- hexas %>%
    filter(ecoregion %in% eco_sim) %>%
    mutate(    
        eco_name = paste0('eco_', ecoregion), # to match design name
        mdcaty  = sum(Stratdsgn) * p/sum(p),
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )
    
# coordinates in matrix format for MBHdesign
coord_mt <- sampleFrame %>%
    st_coordinates()

legacySites <- sampleFrame %>%
    filter(ET_Index %in% subset(selected_hexas, sampled == 1)$ET_Index) %>%
    st_coordinates()


sampleFrame$adj_p <- MBHdesign::alterInclProbs(
    legacy.sites = legacySites,
    potential.sites = coord_mt,
    inclusion.probs = sampleFrame$mdcaty,
    sigma = mean(best_bufferSize$best_bs) * 1000
)

# run GRTS    
out_sample <- spsurvey::grts(
    sframe = sampleFrame,
    n_base = Stratdsgn,
    stratum_var = 'eco_name',
    aux_var = 'adj_p'
)
```


```{r,echo=FALSE,fig.align='center',fig.width=12,fig.height=7,warning=FALSE}
for(eco in eco_sim)
{
    eco_limits <- hexas %>%
        filter(ecoregion == eco) %>%
        st_union()
    
    p1 <- ggplot() +
        geom_sf(data = eco_limits) +
        geom_sf(data = subset(selected_hexas, ecoregion == eco & sampled == 0), fill = 'lightblue', col = 'lightblue') +
        geom_sf(data = subset(selected_hexas, ecoregion == eco & sampled == 1), fill = 'red', col = 'red') +
        ggtitle("**<span style='color:red'>Sampled</span>** (1%) and **<span style='color:lightblue'>non-sampled</span>** (1%) hexagons year 1") +
        theme(plot.title = element_markdown())

    p2 <- ggplot() +
        geom_sf(data = eco_limits) +
        geom_sf(data = st_union(st_buffer(subset(selected_hexas, ecoregion == eco & sampled == 1), dist = mean(best_bufferSize$best_bs) * 1000)), fill = rgb(.85, 0, 0, 0.05), color = 'transparent') +
        geom_sf(data = subset(selected_hexas, ecoregion == eco & sampled == 0), fill = 'lightblue', col = 'lightblue') +
        geom_sf(data = subset(hexas, ecoregion == eco  & ET_Index %in% out_sample$sites_base$ET_Index), fill = 'blue', col = 'blue') +
        ggtitle("**<span style='color:blue;'>Selected</span>** (1%) hexagons year 2") +
        theme(plot.title = element_markdown())

    print(
        ggarrange(p1, p2, nrow = 1) %>%
            annotate_figure(
                top = gridtext::richtext_grob(
                    paste0(
                        '<br><br>**<span style="font-size:14pt;">Ecoregion = ', eco, '</span>**'
                    )        
                )
            )
    )
}
```
