---
title: "How the Nearest Neighbour Method selects over samples"
author: Will Vieira
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Setup

```{r,warning=F,message=F}
library(leaflet)
library(mapview)
library(tidyverse)
library(sf)
library(spatstat)

set.seed(0.0)

hexas <- readRDS('../data/hexa_complete.RDS')

# First filter for hexagons that have at least 20% of sampleable pixel habitats (non NAs)
hexas <- subset(hexas, propNA <= 0.8)

# calculate p
hexas$p <- (hexas$hab_prob * hexas$cost_prob) / sum(hexas$hab_prob * hexas$cost_prob)

# remove zeros
hexas <- subset(hexas, p != 0)

```



```{r}
# Get sample size for each ecoregion
sampleSize <- readRDS('../data/nbHexa_ecoregion.RDS')

# Sample 2% of available hexagons
N_tmp <- sampleSize$hexagonProp80 * 0.02

# Keep only North and South part of splited ecoregions
eco_ignore <- which(sampleSize$ecoregion %in% gsub('N', '', grep('N', sampleSize$ecoregion, value = TRUE)))

# Sample size as a function of ecoregion size
totalN <- sum(N_tmp[-eco_ignore])
sampleSize$N <- unclass(round(totalN * sampleSize$sizeProp, 0))

# Remove ecoregions with N == 0
sampleSize <- subset(sampleSize, N != 0)
```


## Run GRTS

```{r}
# number of replications for each simulation
nb_rep <- 30

# Ecoregions
eco_sim <- c(
    '7', '28', '30', '31', '47', '48', '49', '73', '77', '78', '86',
    '72', '74', '75', '76',
    #'46',
    '96N', #'96', '96S',
    '99',
    '100N', #'100', '100S',
    '101N', #'101', '101S',
    '102',
    '103N', #'103', '103S',
    '117',
    '216',
    '217N'#, '217', '217S'
)
```


```{r}
# Sample size by stratum (ecoregion)
Stratdsgn  <- c()
for(eco in eco_sim)
    Stratdsgn <- append(
        Stratdsgn,
        setNames(
            subset(sampleSize, ecoregion == eco)$N,
            paste0('eco_', eco)
        )
    )

# Sample frame for specific ecoregions
sample_frame <- subset(hexas, ecoregion %in% eco_sim)

# rename ecoregion name to match design name
sample_frame$eco_name <- paste0('eco_', sample_frame$ecoregion)

# Keep only centroid of hexagon for GRTS
sample_frame$geometry <- sample_frame %>%
    sf::st_centroid() %>%
    sf::st_geometry()


# calculate inclusion probability given total sample size
sample_frame$mdcaty <- sum(subset(sampleSize, ecoregion %in% eco_sim)$N) * sample_frame$p/sum(sample_frame$p)
```


```{r}
# list to save nb_rep simulations   
out_sim <- list()

# run GRTS
for (i in 1:nb_rep)
{
    out_sample <- spsurvey::grts(
        sframe = sample_frame,
        n_base = Stratdsgn,
        stratum_var = 'eco_name',
        aux_var = 'mdcaty',
        n_near = 1
    )

    out_sim[[i]] <- c(
        setNames(
            out_sample$sites_base$ET_Index,
            rep('main', nrow(out_sample$sites_base))
        ),
        setNames(
            out_sample$sites_near$ET_Index,
            rep('over', nrow(out_sample$sites_near))
        )
    )
}
```


## Extract inclusion probability of the 6 neighbour hexagons

```{r}
incProb <- data.frame()

for(i in 1:nb_rep)
{
    # get hexagons ID selected for main
    hexas_i_main <- out_sim[[i]][which(names(out_sim[[i]]) == 'main')]

    # extract neigbours inside buffer of 5.5Km around centroid
    neighbours <- st_intersects(
        st_buffer(
            sample_frame[match(hexas_i_main, sample_frame$ET_Index), ],
            dist = 5500
        ),
        sample_frame
    )
    # Transform list to data.frame
    incProb <- rbind(
        incProb,
        data.frame(
            rep = i,
            ET_Index_main = rep(hexas_i_main, lengths(neighbours)),
            ET_Index_near = sample_frame$ET_Index[unlist(neighbours)],
            p = sample_frame$p[unlist(neighbours)]
        )
    )
}

# remove the target hexagon from over column
incProb <- incProb[!incProb$ET_Index_main == incProb$ET_Index_near, ]

# define which neighbour hexagon was selected
incProb$selected <- 0

for(i in 1:nb_rep)
{
    # get hexagons ID selected for main
    hexas_i_main <- out_sim[[i]][which(names(out_sim[[i]]) == 'main')]
    hexas_i_over <- out_sim[[i]][which(names(out_sim[[i]]) == 'over')]

    # I do need a second loop to avoid the same hexagon being a neighb twice     
    for(j in 1:length(hexas_i_main))
    {
        row_ij <- which(
                incProb$rep == i &
                incProb$ET_Index_main == hexas_i_main[j] &
                incProb$ET_Index_near == hexas_i_over[j]
            )
        incProb$selected[row_ij] <- 1
    }
}
```


## Analysis

### Does the selected "over" hexagon have the highest inclusion probability among the 6 neighboring hexagons? 

We extracted the inclusion probability $p$ from the 6 neighboring hexagons around the "main" hexagon, sorted these $p$ in ascending order, and classified the ranking position of the selected "over" hexagon. For instance, if the selected "over" hexagon has the highest probability $p$ among the 6 neighbours, its ranking position will be 1.

```{r,message=F, warning=F, echo=F, fig.align='center', fig.width=8}
incProb %>%
    group_by(rep, ET_Index_main) %>%
    mutate(
        sel_total = sum(selected)
    ) %>%
    group_by(rep, ET_Index_main) %>%
    filter(sel_total == 1) %>%
    group_by(rep, ET_Index_main) %>%
    summarise(
        p_pos = which(p[selected == 1] == sort(p))/length(p),
        p_selected = p[selected == 1],
        max_p = max(p),
        min_p = min(p)
    ) %>%
    ggplot(aes(p_pos)) +
        geom_histogram() +
        xlab('Ranking position') +
        ggtitle('Distribution of ranking position for all selected "over" hexagons', subtitle = paste0('Ecoregions: ', paste0(eco_sim, collapse = ', '), '\n(30 replications)'))
```

```{r, echo=F}
# position_dist %>%
#     ggplot(aes(max_p, min_p, color = p_pos)) +
#         geom_point(alpha = .2) +
#         xlim(0, 0.0002) + 
#         ylim(0, 0.0001)
```


## Example

Ecoregion 101 nord:

```{r, fig.width=9.5, fig.height=8,echo=F}
hex_101 <- subset(hexas, ecoregion %in% c('101N'))

hex_101$selection <- NA
hex_101$selection[
    hex_101$ET_Index %in% out_sim[[1]][which(names(out_sim[[1]]) == 'main')]
] <- 'main'
hex_101$selection[
    hex_101$ET_Index %in% out_sim[[1]][which(names(out_sim[[1]]) == 'over')]
] <- 'over'

mapview(hex_101[, c('ET_Index', 'selection', 'p')], zcol = 'selection')
```
