---
title: "How to account for historical (Legacy) sites using the GRTS function"
author: Will Vieira
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r libraries,warning=F,message=F}
library(leaflet)
library(mapview)
library(tidyverse)
library(sf)
library(spatstat)
library(kableExtra)
library(spsurvey)
library(MBHdesign)

set.seed(0.0)

hexas <- readRDS('../data/hexa_complete.RDS')

# First filter for hexagons that have at least 20% of sampleable pixel habitats (non NAs)
hexas <- subset(hexas, propNA <= 0.8)

# calculate p
hexas$p <- (hexas$hab_prob * hexas$cost_prob) / sum(hexas$hab_prob * hexas$cost_prob)

# remove zeros
hexas <- subset(hexas, p != 0)

```



```{r sampleSize}
# Get sample size for each ecoregion
sampleSize <- readRDS('../data/nbHexa_ecoregion.RDS')

# Sample 2% of available hexagons
N_tmp <- sampleSize$hexagonProp80 * 0.02

# Keep only North and South part of splited ecoregions
eco_ignore <- which(sampleSize$ecoregion %in% gsub('N', '', grep('N', sampleSize$ecoregion, value = TRUE)))

# Sample size as a function of ecoregion size
totalN <- sum(N_tmp[-eco_ignore])
sampleSize$N <- unclass(round(totalN * sampleSize$sizeProp, 0))

# Remove ecoregions with N == 0
sampleSize <- subset(sampleSize, N != 0)


# number of replications for each simulation
nb_rep <- 10

# Ecoregions
eco_sim <- c(
    '7', '28', '30', '31', '47', '48', '49', '73', '77', '78', '86',
    '72', '74', '75', '76',
    #'46',
    '96N', '96S',
    '99',
    '100N', '100S',
    '101N', '101S',
    '102',
    '103N', '103S',
    '117',
    '216',
    '217N', '217S'
)


# filter for ecoregions with at least 1 legacy site
eco_sim <- hexas %>%
    st_drop_geometry() %>%
    filter(ecoregion %in% eco_sim) %>%
    group_by(ecoregion) %>%
    summarise(total_legacy = sum(legacySite)) %>%
    filter(total_legacy > 0) %>%
    pull(ecoregion)
    
```


## Run GRTS


```{r GRTS setup}
# Sample size by stratum (ecoregion)
Stratdsgn_noLegacy  <- c()
for(eco in eco_sim)
    Stratdsgn_noLegacy <- append(
        Stratdsgn_noLegacy,
        setNames(
            subset(sampleSize, ecoregion == eco)$N,
            paste0('eco_', eco)
        )
    )

# Sample size by stratum (ecoregion) + nb legacy Site (required by GRTS)
Stratdsgn_legacy  <- c()
for(eco in eco_sim)
    Stratdsgn_legacy <- append(
        Stratdsgn_legacy,
        setNames(
            subset(sampleSize, ecoregion == eco)$N +
            sum(subset(hexas, ecoregion == eco)$legacySite > 0),
            paste0('eco_', eco)
        )
    )


# Sample frame for specific ecoregions
sampleFrame_legacy <- sampleFrame_noLegacy <- subset(hexas, ecoregion %in% eco_sim)

sampleFrame_noLegacy <- sampleFrame_noLegacy %>%
    mutate(
        eco_name = paste0('eco_', ecoregion), # to match design name
        mdcaty = sum(subset(sampleSize, ecoregion %in% eco_sim)$N) *
        p/sum(p), # p given total sample size
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )


sampleFrame_legacy <- sampleFrame_legacy %>%
    mutate(
        eco_name = paste0('eco_', ecoregion), # to match design name
        mdcaty = (sum(subset(hexas, ecoregion %in% eco_sim)$legacySite > 0) + sum(subset(sampleSize, ecoregion %in% eco_sim)$N)) *
        p/sum(p), # p given total sample size
        legacy_var = ifelse(legacySite == 0, NA, legacySite),
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )
```

```{r run GRTS}
# list to save nb_rep simulations   
out_legacy <- out_noLegacy <- list()

# run GRTS
for (i in 1:nb_rep)
{
    out_sample_noLegacy <- spsurvey::grts(
        sframe = sampleFrame_noLegacy,
        n_base = Stratdsgn_noLegacy,
        stratum_var = 'eco_name',
        aux_var = 'mdcaty',
        n_near = 1
    )

    out_sample_legacy <- spsurvey::grts(
        sframe = sampleFrame_legacy,
        n_base = Stratdsgn_legacy,
        stratum_var = 'eco_name',
        aux_var = 'mdcaty',
        n_near = 1,
        legacy_var = 'legacy_var'
    )

    out_noLegacy[[i]] <- list(
        main = out_sample_noLegacy$sites_base$ET_Index,
        over = out_sample_noLegacy$sites_near
    )
    out_legacy[[i]] <- list(
        main = out_sample_legacy$sites_base$ET_Index,
        over = out_sample_legacy$sites_near
    )
}
```

## Analysis

<!--
## How many legacy sites are selected on average when using the `legacy_var` option from GRTS?

Compared when no legacy information is used (`noLegacy` column)
--->

```{r analysis1,echo=F,eval=F}
nbLegacy_dt <- data.frame()

for(eco in eco_sim)
{
    for(i in 1:nb_rep)
    {
        nb_legacy <- hexas %>%
            st_drop_geometry() %>%
            filter(ET_Index %in% out_legacy[[i]]$main) %>%
            filter(ecoregion == eco) %>%
            summarise(nb = sum(legacySite > 0)) %>%
            pull(nb)
        
        nb_noLegacy <- hexas %>%
            st_drop_geometry() %>%
            filter(ET_Index %in% out_noLegacy[[i]]$main) %>%
            filter(ecoregion == eco) %>%
            summarise(nb = sum(legacySite > 0)) %>%
            pull(nb)

        nbLegacy_dt <- rbind(
            nbLegacy_dt,
            data.frame(
                ecoregion = eco,
                Rep = i,
                nb_legacy = nb_legacy,
                nb_noLegacy = nb_noLegacy
            )
        )
    }
}
```


```{r plot1,echo=F, eval=FALSE}
nbLegacy_dt %>%
    group_by(ecoregion) %>%
    summarise(
        legacy_var = mean(nb_legacy),
        noLegacy = mean(nb_noLegacy)
    ) %>%
    kbl() %>%
    kable_paper("hover", full_width = F)
```


### Using the `legacy_var` option on GRTS will increase the average distance between selected sites and legacy sites?

For each replication, I computed the average euclidean distance between the selected sites and the legacy sites for two simulations: (i) with and (ii) without legacy information provided to the GRTS function.
I expect the average distance of the `legacy_var` simulation to be smaller than the `noLegacy` simulation.

```{r analysis2,echo=F}
dist_dt <- data.frame()
dist_min <- data.frame()

for(eco in eco_sim)
{
    legacy <- hexas %>%
        filter(ecoregion == eco & legacySite > 0)
    legacy_len <- nrow(legacy)

    for(rep in 1:nb_rep)
    {
        selected_legacy <- hexas %>% 
            filter(ET_Index %in% out_legacy[[rep]]$main) %>%
            filter(ecoregion == eco)
        selected_noLegacy <- hexas %>% 
            filter(ET_Index %in% out_noLegacy[[rep]]$main) %>%
            filter(ecoregion == eco)
        
        Dist_mt_legacy <- st_distance(selected_legacy, legacy)
        Dist_mt_noLegacy <- st_distance(selected_noLegacy, legacy)

        dist_dt <- rbind(
            dist_dt,
            data.frame(
                ecoregion = eco,
                Rep = rep,
                sim = c(
                    rep('legacy_var', nrow(selected_legacy) * legacy_len),
                    rep('noLegacy', nrow(selected_noLegacy) * legacy_len)
                    ),
                dist = c(
                    as.vector(Dist_mt_legacy),
                    as.vector(Dist_mt_noLegacy)
                )
            )
        )

        dist_min <- rbind(
            dist_min,
            data.frame(
                ecoregion = eco,
                Rep = rep,
                sim = c(
                    rep('legacy', nrow(selected_legacy)),
                    rep('noLegacy', nrow(selected_noLegacy))
                    ),
                dist = c(
                    apply(Dist_mt_legacy, 1, min),
                    apply(Dist_mt_noLegacy, 1, min)
                )
            )
        )
    }
}
```


```{r plot2,echo=F,fig.width=12,fig.height=10}
dist_dt %>%
    mutate(sim = as.factor(sim)) %>%
    ggplot(aes(y = dist, x = sim, color = sim, fill = sim)) +
        facet_wrap(~ecoregion, scales = 'free') +
        geom_boxplot(alpha = 0.5) +
        xlab('') +
        ylab('Average distance between selected sites and legacy sites (m)') +
        theme_classic() +
        theme(
            legend.position = 'none'
        )
```




### What if we use the distance from the closest legacy site?

As there was no difference between the simulations, I computed the distance for each selected hexagon to its closest legacy site.
This may avoid the loss of signal when averaging the distance of a selected site with all legacy sites distributed over the ecoregion.


```{r plot3,echo=F,fig.width=12,fig.height=10}
dist_min %>%
    mutate(sim = as.factor(sim)) %>%
    ggplot(aes(y = dist, x = sim, color = sim, fill = sim)) +
        facet_wrap(~ecoregion, scales = 'free') +
        geom_boxplot(alpha = 0.5) +
        xlab('') +
        ylab('Distance between a selected site and its closest legacy sites (m)') +
        theme_classic() +
        theme(
            legend.position = 'none'
        )

# dist_min %>%
#     mutate(sim = as.factor(sim)) %>%
#     ggplot(aes(y = dist, x = sim, color = sim, fill = sim)) +
#         facet_grid(ecoregion ~ Rep, scales = 'free') +
#         geom_boxplot(alpha = 0.5) +
#         xlab('') +
#         theme_classic() +
#         theme(
#             legend.position = 'none'
#         )
```





## Run GRTS with altered inclusion probability using the `MBHdesign` R package

As the native legacy functionality in the GRTS does not account for the distance between potential sites and legacy sites, here I will use the `MBHdesign` R package to change the inclusion probability of potential sites according to their distance to legacy sites.
The inclusion probability of potential sites is changed as a function of their distance to nearby legacy sites, according to the negative exponential distribution with parameter $\sigma$.

```{r,echo=FALSE,fig.align='center',fig.width=5,fig.height=3.5}
par(mar = c(2.2, 2.2, 1.2, 0.8), oma = c(0, 0, .5, 0), mgp = c(1.2, 0.2, 0), tck = -.008, cex = 0.8)
# exp function
Exp = function(dist, sigma) return ( exp(-dist^2/sigma^2) )
# Diff parameter values
sigmas = c(1000, 4000, 6000, 10000, 15000, 20000)
# viz
plot(0, pch = '', xlim = c(0, 40000), ylim = c(0, 1), type = 'l', xlab = 'distance from legacy site (m)', ylab = 'Legacy effect')
for(i in sigmas) 
    lines(Exp(1:40000, sigma = i), col = which(i == sigmas))  
legend('topright', legend = sigmas, col = 1:length(sigmas), lty = 1, bty = 'n')
```

### Example of altered inclusion probability for different values of $\sigma$

Example from the `MBHdesign` R package vignette ([link](https://cran.r-project.org/web/packages/MBHdesign/vignettes/MBHdesign.pdf)).

```{r example1,fig.width=10,fig.height=8,warning=FALSE}
#number of samples
n <- 10
#number of points to sample from
N <- 100^2
#the sampling grid (offset so that the edge locations have same area)
offsetX <- 1/(2*sqrt( N))
my.seq <- seq(from=offsetX, to=100000-offsetX, length=sqrt(N))
X <- expand.grid( my.seq, my.seq)
#the legacy sites (three of them)
set.seed(10)
legacySites <- matrix(sample(1:100000, 6, replace = FALSE), ncol = 2, byrow = TRUE)
#names can be useful
colnames(X) <- colnames( legacySites) <- c("X1","X2")

#non-uniform inclusion probabilities
inclProbs <- 1-exp(-X[,1]/100000)
#scaling to enforce summation to n
inclProbs <- n * inclProbs / sum( inclProbs)

# adj according to different values of sigma
sigmas <- c(2000, 10000, 40000)

incl_ls <- list('Unadjusted inclusion Probabilities' = inclProbs)
for(sgm in sigmas)
    incl_ls[[paste0('Adjusted inclusion probabilities (σ = ', sgm, ')')]] <- MBHdesign::alterInclProbs(
        legacy.sites = legacySites, 
        potential.sites = X,
        inclusion.probs = inclProbs,
        sigma = sgm
    ) 

par(mfrow = c(2, 2), mar = c(0.5, 0.5, 1.5, 0.5), oma = c(1.5, 1.5, .5, 0), mgp = c(1.2, 0.2, 0), tck = -.008, cex = 0.8)
for(i in 1:length(incl_ls)) {
    plot(X, pch = 19, cex = 1.2, col = viridis::viridis(N)[rank(incl_ls[[i]])], 
    xlab = '', ylab = '', xaxt = 'n', yaxt = 'n')
    axis(1, labels = ifelse(i %in% c(3, 4), T, F))
    axis(2, labels = ifelse(i %in% c(1, 3), T, F))
    mtext(names(incl_ls)[i], 3, line = 0) 
    points(legacySites, pch = 21, bg = 'grey', cex = 1.5)
}
```


### Run GRTS with alterd inclusion probabilities

```{r run GRTS2,warning=FALSE}
sampleFrame_legacy <- subset(hexas, ecoregion %in% eco_sim)

sampleFrame_legacy <- sampleFrame_legacy %>%
    mutate(
        eco_name = paste0('eco_', ecoregion), # to match design name
        mdcaty = sum(subset(sampleSize, ecoregion %in% eco_sim)$N) *
        p/sum(p), # p given total sample size
        legacy_var = ifelse(legacySite == 0, NA, legacySite),
        geometry = sf::st_geometry(sf::st_centroid(geometry))
    )


# coordinates in matrix format for MBHdesign
coord_mt <- sampleFrame_legacy %>%
    st_coordinates()

legacySites <- sampleFrame_legacy %>%
    filter(legacySite > 0) %>%
    st_coordinates()

sigmas <- c(1000, 5000, 10000, 15000, 20000)
for(sgm in 1:length(sigmas))
    sampleFrame_legacy[[paste0('pAdj_', sgm)]] <- 
        MBHdesign::alterInclProbs(
            legacy.sites = legacySites,
            potential.sites = coord_mt,
            inclusion.probs = sampleFrame_legacy$mdcaty,
            sigma = sigmas[sgm]
        )

```

#### Ecoregion 101 as an example

Normalized inclusion probabilities for different values of $\sigma$.
All hexagons with one or more legacy sites are colored in red.

```{r example2,echo=FALSE, fig.align='center', fig.width=12,fig.height=15,warning=FALSE}
hexas %>%
    filter(ecoregion %in% eco_sim) %>%
    bind_cols(st_drop_geometry(sampleFrame_legacy[, c('ET_Index', 'mdcaty', paste0('pAdj_', 1:length(sigmas)))])) %>%
    filter(ecoregion %in% c('101S', '101N') & mdcaty < 0.1) %>%
    mutate(
        sim1 = (mdcaty - min(mdcaty))/(max(mdcaty) - min(mdcaty)),
        sim2 = (pAdj_1 - min(pAdj_1))/(max(pAdj_1) - min(pAdj_1)),
        sim3 = (pAdj_2 - min(pAdj_2))/(max(pAdj_2) - min(pAdj_2)),
        sim4 = (pAdj_3 - min(pAdj_3))/(max(pAdj_3) - min(pAdj_3)),
        sim5 = (pAdj_4 - min(pAdj_4))/(max(pAdj_4) - min(pAdj_4)),
        sim6 = (pAdj_5 - min(pAdj_5))/(max(pAdj_5) - min(pAdj_5))
    ) %>%
    select(ecoregion, starts_with('sim')) %>%
    pivot_longer(cols = starts_with('sim')) %>%
    ggplot() +
        geom_sf(aes(fill = value), lwd = 0) +
        viridis::scale_fill_viridis(option = "D") +
        geom_sf(data = subset(sampleFrame_legacy, ecoregion %in% c('101S', '101N') & legacySite > 0), color = 'red', alpha = 0.7, size = 0.15) +
        facet_wrap(
            ~name,
            labeller = labeller(
                name = setNames(
                    c('Unadjusted inclusion Probabilities', paste0('σ = ', sigmas)),
                    paste0('sim', 1:6)
                )
            )
        ) + 
        theme(legend.position = 'none')
    
```


### Run GRTS

```{r run GRTS3,eval=F}
out_sigmas <- list('Unadjusted' = out_legacy)

for(sgm in 1:length(sigmas))
{
    out_ls <- list()
    # run GRTS
    for (i in 1:nb_rep)
    {
        out_sample_legacy <- spsurvey::grts(
            sframe = sampleFrame_legacy,
            n_base = Stratdsgn_legacy,
            stratum_var = 'eco_name',
            aux_var = paste0('pAdj_', sgm),
            n_near = 1
        )

        out_ls[[i]] <- list(
            main = out_sample_legacy$sites_base$ET_Index,
            over = out_sample_legacy$sites_near
        )
    }

    out_sigmas[[paste0('sigma_', sigmas[sgm])]] <- out_ls
}
```




## Analysis

### Will altered inclusion probabilities increase the average distance between selected sites and legacy sites?

For each replication, I computed the average euclidean distance between the selected sites and the legacy sites for two simulations: (i) Unadjusted vs (ii)adjusted inclusion probabilities in function of the distance to legacy sites.

```{r analyses2,echo=F}
# dist_dt <- data.frame()
# dist_min <- data.frame()

# for(eco in eco_sim)
# {
#     legacy <- hexas %>%
#         filter(ecoregion == eco & legacySite > 0)

#     if(dim(legacy)[1] > 0) {
#         for(sgm in 1:length(out_sigmas))
#         {
#             for(Rep in 1:nb_rep)
#             {
#                 selected_hexas <- hexas %>% 
#                     filter(ET_Index %in% out_sigmas[[sgm]][[Rep]]$main) %>%
#                     filter(ecoregion == eco)
                
#                 Dist_mt <- st_distance(selected_hexas, legacy)
                
#                 dist_dt <- rbind(
#                     dist_dt,
#                     data.frame(
#                         ecoregion = eco,
#                         sigma = sgm,
#                         Rep = Rep,
#                         dist = as.vector(Dist_mt)
#                     )
#                 )

#                 dist_min <- rbind(
#                     dist_min,
#                     data.frame(
#                         ecoregion = eco,
#                         sigma = sgm,
#                         Rep = Rep,
#                         dist = apply(Dist_mt, 1, min)
#                     )
#                 )
#             }
#         }
#     }
# }

dist_dt <- readRDS('../dist_dt.RDS')
dist_min <- readRDS('../dist_min.RDS')
```


```{r plot4,echo=F,fig.width=12,fig.height=12}
dist_dt %>%
    mutate(sigma = as.factor(sigma)) %>%
    ggplot(aes(y = dist, x = sigma, fill = sigma, color = sigma)) +
        scale_fill_hue(
            l = 45,
            labels = c('Unadjusted', paste0('sigma = ', sigmas))
        ) +
        scale_color_hue(
            l = 45,
            labels = c('Unadjusted', paste0('sigma = ', sigmas))
        ) +
        facet_wrap(~ecoregion, scales = 'free') +
        geom_boxplot(alpha = 0.5) +
        xlab('') +
        ylab('Average distance between selected sites and legacy sites (m)') +
        theme_classic() +
        theme(
            legend.position = 'bottom'
        )
```

### Distance to the closest legacy site

```{r plot5,echo=F,fig.width=12,fig.height=12}
dist_min %>%
    mutate(sigma = as.factor(sigma)) %>%
    ggplot(aes(y = dist, x = sigma, fill = sigma, color = sigma)) +
        scale_fill_hue(
            l = 45,
            labels = c('Unadjusted', paste0('sigma = ', sigmas))
        ) +
        scale_color_hue(
            l = 45,
            labels = c('Unadjusted', paste0('sigma = ', sigmas))
        ) +
        facet_wrap(~ecoregion, scales = 'free') +
        geom_boxplot(alpha = 0.5) +
        xlab('') +
        ylab('Distance between a selected site and its closest legacy sites (m)') +
        theme_classic() +
        theme(
            legend.position = 'bottom'
        )
```














